---
title: "Homework 2"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{unicode-math}
date: "2023-09-19"
---
Problem 1

To avoid an error, dummy variable trap we keep one category as a reference group, remaining ones represent the change from the reference. That is why we will have n-1 number of dummy variables for independent variable with ğ‘› categories. 

Problem 2 

x is a dummy variable changing values between 0 and 1. 
$ğ›½1$ represents the change in the predicted value of y E(log(y)) when x changes.

Problem 3
```{r}
# Loading the data
data <- read.csv("houseprice.csv")
#str(data)

# Fitting the regression model
model <- lm(price ~ sqft + age + sqft:age, data=data)

# Summarizing the model
summary(model)

# Extracting the coefficient for the interaction term
gamma <- coef(model)["sqft:age"]

# Print the estimated value of gamma
cat("The estimated value of gamma is:", gamma)
```
Gamma $ğ›¾$ (interaction term coefficient) represents the effect of the interaction between the size of the house and its age on the selling price.

The estimated value of gamma $\hat{ğ›¾}$ means that when we change the age by one unit and keep all the other variables constant, it shows the relationship between sqft and price.

```{r}
# Finding the minimum and maximum values
min_sqft <- min(data$sqft)
max_sqft <- max(data$sqft)

# Calculating the marginal effect for the smallest house
marginal_effect_1 <- coef(model)["age"] + gamma * min_sqft

# Calculate the marginal effect for the largest house
marginal_effect_2 <- coef(model)["age"] + gamma * max_sqft

# Print the results
cat("Marginal Effect for Smallest House:", marginal_effect_1)
cat("Marginal Effect for Largest House:", marginal_effect_2)
```
```{r}
# Calculating the marginal effect
age <- 20 
marginal_effect <- coef(model)["sqft"] + gamma * age

cat("Marginal Effect of sqft on price for a 20-year-old house:", marginal_effect)
#how the price of 20 years old house is changing when 1 additional sqft is added.
```
We can interpret it as the change in the predicted price for a 1-unit increase in sqft while having the age fixed at 20 years.
```{r}
library(car)

# Performing the F-test
f_test <- linearHypothesis(model, c("sqft = 0", "age = 0"))
summary(f_test)

# Extracting the p-value
p <- f_test$`Pr(>F)`

# Setting the significance level
a <- 0.05

# Checking if the p-value is less than the significance level
result <- ifelse(p < a, "Reject", "Fail to reject")
result
```
That means at least one of $ğ›½2$ or $ğ›½3$ is significant.

Problem 4
```{r}
#1
library(ggplot2)

# Load the data
data <- read.csv("fullmoon.csv")

# Examine the histogram of cases
ggplot(data, aes(x = cases)) +
  geom_histogram(fill = "blue", color = "black") +
  labs(title = "Histogram of Cases", x = "Cases", y = "Frequency")

# Create the variable ln(cases)
data$ln_cases <- log(data$cases)

# Examine the histogram of ln(Cases)
ggplot(data, aes(x = ln_cases)) +
  geom_histogram(fill = "turquoise", color = "black") +
  labs(title = "Histogram of ln(Cases)", x = "ln(Cases)", y = "Frequency")
```
In the histogram of cases (without logarithm) we see that it is right skewed.
 
The histogram of cases with logarithm is much closer to Normal distribution.

So, by taking the natural logarithm of a variable we can make the distribution more symmetric and reduce the impact of extreme values.
```{r}
#2
model_1 <- lm(log(cases) ~ time + holiday + friday + saturday + fullmoon + newmoon, data = data)
summary(model_1)
```
Based on the regression results, it is not possible to conclude that there is a statistically significant effect of a full moon on emergency room cases.

However, we see that the variables time, holiday, friday, and saturday (saturday is highly significant) are statistically significant.
The time variable is positively associated with log(cases), meaning that as time increases, log(cases) tends to increase.
Holiday, Friday, and Saturday variables are also positively associated with log(cases), indicating that these factors tend to increase log(cases) compared to non-holiday days and other weekdays. 

The fullmoon and  the newmoon are not statistically significant. That means presence does not have any significant effect on the number of cases.
```{r}
#3
```
The coefficient for the fullmoon is not statistically significant because the p-value = 0.52211. The presence of it does not have any significant effect on the number of cases log(cases) when controlling for other variables in the model.
```{r}
#4
library(car)

# Performing the F-test
f_test <- linearHypothesis(model_1, c("friday = 0", "saturday = 0"))
summary(f_test)

# Extracting the p-value
p <- f_test$`Pr(>F)`

# Setting the significance level
a <- 0.95

# Checking if the p-value is less than the significance level
result <- ifelse(p < a, "Reject", "Fail to reject")
result
```
That means at least one of friday or saturday is significant.
```{r}
#5
model_2 <- lm(log(cases) ~ time + holiday + saturday + fullmoon + newmoon, data = data)
summary(model_2)
```
After dropping friday variable from the model, we see that p-value of fullmoon changed to 0.55944, but the difference is really small even after removing friday. So, the effect of a fullmoon remains statistically insignificant.
```{r}
#6
data$interval_time <- cut(data$time, breaks = c(0, 100, 200, Inf), labels = c("0-100 days", "101-200 days", "200+ days")) 

model_3 <- lm(log(cases) ~ interval_time + holiday + saturday + fullmoon + newmoon, data = data)
summary(model_3)
```
R has automatically treated "0-100" as a reference group, and that is why it does not display "0-100" coefficients. It provides coefficients for the other groups in relation to the reference one. 

In the regression model, the coefficients of the interval_time variable represent the estimated effects of different time intervals on log(cases). We see that "101-200" interval has a stronger and more statistically significant effect compared to "200+" interval. So, a longer time period within the range of "101-200" tends to be associated with a higher number of cases, and this effect is statistically supported.
