---
title: 'Homework 8'
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{unicode-math}
date: "2023-12-08"
---

```{r}
# Data 
data_A <- c(62, 65, 58, 60, 64, 61, 63, 59, 66, 62)
data_B <- c(68, 72, 70, 71, 69, 73, 67, 70, 75, 74)

# Creating histograms
par(mfrow = c(1, 2)) 

# Histogram for Fertilizer A
hist(data_A, breaks = 5, col = "darkgray", 
     main = "Fertilizer A Plant Heights", xlab = "Height (cm)", 
     ylab = "Frequency")

# Histogram for Fertilizer B
hist(data_B, breaks = 5, col = "darkgray", 
     main = "Fertilizer B Plant Heights", 
     xlab = "Height (cm)")
```
```{r}
# Performing a t-test
t_test_result <- t.test(data_A, data_B)
t_test_result
```

The negative t-statistic indicates that plants treated with Fertilizer B are, on average, taller than those treated with Fertilizer A.

The extremely low p-value suggests that this difference is statistically significant and not due to random chance.

The 95% confidence interval for the difference in means does not include zero, reinforcing the conclusion that there is a significant difference.

The sample means show that plants treated with Fertilizer B are, on average, about 8.9 cm taller than those treated with Fertilizer A.

```{r}
#install.packages("BayesFactor")
library(BayesFactor)

# Performing Bayesian t-test
bayes_test_result <- ttestBF(x = data_A, y = data_B)
bayes_test_result

```
Bayes Factor is 22550.34.

The Bayes Factor significantly favors the alternative hypothesis (true difference in means is not equal to zero) over the null hypothesis.

A Bayes Factor greater than 1 indicates evidence against the null hypothesis, and in this case, the value is extremely high, providing strong evidence for a difference in mean heights between the two groups.



Overall Conclusion:

Both the Welch Two Sample t-test and the Bayesian analysis provide strong evidence that there is a significant difference in the mean height of plants treated with the two different fertilizers, with Fertilizer B leading to taller plants on average compared to Fertilizer A. This conclusion is robust across both frequentist and Bayesian statistical methods.
